Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-12 introduction.tex
a couple of comments in section 1

\MS{Here, I would just say that $B(a,b)$ denotes the beta function $B^-(x,a,b)$ the lower incomplete beta function etc. and put/refer to the integral representations in the appendix, as already done for the Meijer-G function.}
%\subsubsection{Special functions}
We let $B(a,b)$ denote the beta-function and $B^-(x,a,b)$ the lower incomplete beta-function. We also write $\Gamma(z)$ for the Gamma function, $\Gamma^+(q,z)$ for the incomplete Gamma function and define $\Gamma^\ast(q,z) := \Gamma^+(1 + q, z) + \Gamma^+(q, z)$. Finally, we write $U(a,b,z)$ for the hypergeometric U-function (also called Tricomi's confluent hypergeometric function), which for $a,b,z\in \mathbb{C}$, $b \not \in \mathbb{Z}_{\leq 0}$, $\mathrm{Re}(a), \mathrm{Re}(z) >0$ has the integral representation 
\[
	U(a,b,z) = \frac{1}{\Gamma(a)} \int_0^\infty e^{-zt} t^{a-1} (1+t)^{b-a-1} dt,
\] 
see~\cite[p.255 Equation (2)]{erdelyi1953higher}, and let $\MeijerGnew{m}{\ell}{p}{q}{{\bf a}}{{\bf b}}{z}$ denote Meijer's G-Function~\cite{meijer1946gfunction}, see Appendix~\ref{sec:Meijer_G_functions} for more details.

@@

Our first result shows that the local clustering coefficient of the hyperbolic random graph converges in expectation to as constant $c_\infty$, which we can explicitly write down in terms of the special function defined above.

\begin{theorem}[Limit for local clustering coefficient in $G_{\H,n}(\alpha,\nu)$] \label{thm:clustering_coefficient_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$. Then,
\[
	\lim_{n \to \infty} \Exp{\left|c_{\H,n} - c_\infty\right|} = 0,
\]
where $c_\infty$ is defined for $\alpha \ne 1$ as
\begin{align*}
	c_\infty 
	&=\frac{2 + 4 \alpha + 13 \alpha^2 - 34 \alpha^3 - 12\alpha^4 + 24 \alpha^5}
		{16(\alpha-1)^2 \alpha (\alpha+1) (2\alpha+1)} 
		+  \frac{2^{-1 - 4 \alpha}}{(\alpha - 1)^2} \\
	&\hspace{10pt}+ \frac{(\alpha - 1/2) (B(2 \alpha, 2 \alpha + 1) + B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha))}
		{2 (\alpha - 1) (3 \alpha - 1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{2\alpha} \Gamma^\ast( - 2 \alpha, \xi_{\alpha,\nu})}{4(\alpha-1)}
		+ \frac{\xi_{\alpha,\nu}^{2\alpha + 2}\alpha (\alpha - 1/2)^2 \Gamma^\ast(- 2 \alpha - 2, \xi_{\alpha,\nu})}
		{2(\alpha-1)^2} \\
	&\hspace{10pt}- \frac{\xi_{\alpha,\nu}^{2\alpha + 1}\alpha (2\alpha - 1) \Gamma^\ast( - 2 \alpha - 1,\xi_{\alpha,\nu})}{(\alpha-1)}
		- \frac{\xi_{\alpha,\nu}^{6\alpha-2}2^{-4\alpha}(3\alpha - 1)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)^2}\\
	&\hspace{10pt}-\frac{\xi_{\alpha,\nu}^{6\alpha - 2}(\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)} \\
	&\hspace{10pt}- \frac{e^{-\xi_{\alpha,\nu}} \Gamma(2\alpha+1) 
		\left(U(2\alpha+1,1-2\alpha,\xi_{\alpha,\nu}) + U(2\alpha+1,2-2\alpha,\xi_{\alpha,\nu})\right)}{4(\alpha-1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{6\alpha - 2} \Gamma(2\alpha+1)\left( 	
		\MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+2,0}{\xi_{\alpha,\nu}}
		+ \MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+3,0}{\xi_{\alpha,\nu}}\right)}{4(\alpha-1)},
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty &= \frac{575 - 12 \pi^2}{576} + \frac{\eta^4(7 + \pi^2)\Gamma^\ast(-4, \eta)}{4}\\
	&\hspace{10pt}- \frac{1}{2} \int_0^1 (1 - 4z + 3z^3)\log(1-z)(z + \eta)e^{-\eta/z} \dd z\\
	&\hspace{10pt}- \int_0^1 \Li_2(z)(z^3 + \eta z^2) e^{-\eta/z} \dd z,		
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function\footnote{Note that the integrals in the expression for $c_\infty$ for $\alpha = 1$ exists: for the first one note that $1-4z+3z^2=(1-z)(1-3z)$, so the integrand can be bounded by $C(1-z)\log(1-z)$ on $[0,1)$ for some constant $C$, which can be continued continuously to the compact interval $[0,1]$ by noting that the limit for $z \rightarrow 1$ is zero, so the integrand is bounded on a bounded domain and hence, this integral is finite; for the second integral note that $\Li_2(z)$ is bounded by $\Li_2(1)$ on $[0,1]$, which is a series with well-known finite limit, so again the integrand is bounded on a bounded domain and hence the second integral is also finite.}.
\end{theorem}

@@

\begin{theorem}[Limit for local clustering function in $G_{\H,n}(\alpha, \nu)$]
\label{thm:local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $(k_n)_{n \ge 1}$ be any positive sequence, possibly constant, such that $k_n \ge 3$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, 
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{c_\infty(k_n)} - 1\right|} = 0,
\]
where the function $c_\infty(k)$ is defined for $\alpha \ne 1$ as
\begin{align*}
c_\infty(k)  =&\frac{1}{8\alpha (\alpha-1)\Gamma^+(k-2\alpha,\xi)} \left( -\Gamma^+(k - 2 \alpha, \xi) - 2\frac{\alpha (\alpha - 1/2)^2 \xi^{2} \Gamma^+(k - 2 \alpha - 2, \xi)}{(\alpha - 1)} \right. \\ 
&\left.+ 8 \alpha (\alpha - 1/2) \xi \Gamma^+(k - 2 \alpha - 1,\xi) \right.\\ 
&\left.+ 4\xi^{4\alpha - 2} \Gamma^+(k - 6 \alpha + 2, 
      \xi) \left( \frac{2^{ - 4\alpha}(3 \alpha - 1)}{(\alpha - 1)} + (\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha) \right)  \right.\\ 
&\left.+ \xi^{k-2\alpha} \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi) \right. \\ 
&\left.- \xi^{4\alpha-2} \Gamma(2\alpha+1)\MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}  \right)
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty(k) &= \frac{9 \eta^3}{2 k!} 	
		\Gamma^+(k-3,\eta)-\frac{\xi_{\alpha,\nu}^4}{k!}\frac{7+\pi^2}{4}\Gamma^+(k-4,\eta)\\
	&\hspace{10pt}+ \frac{\eta^k}{2k!}\int_0^1 (1-4z+3z^2)\ln(1-z)z^{1-k}e^{-\eta/z}\dd z\\ 
	&\hspace{10pt}+ \frac{\eta^k}{k!}\int_0^1 z^{3-k} \Li_2(z) e^{-\eta/z} \dd z,
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function.
\end{theorem}

@@

\begin{theorem}[Asymptotic behavior of local clustering function limit]\label{thm:asymptotics_average_clustering_P}
Let $\alpha > \frac{1}{2}$, $\nu > 0$. Then,
\[
	\lim_{k \to \infty} \frac{c_\infty(k)}{s_\alpha(k)} 
	= \begin{cases}
			C_\alpha \left(\frac{4\alpha \nu}{\pi\left(2\alpha - 1\right)}\right)^{4\alpha - 2}
			&\mbox{if } \frac{1}{2} < \alpha < \frac{3}{4},\\
			\frac{3 \nu}{\pi} &\mbox{if } \alpha = \frac{3}{4},\\
			\frac{8\alpha \nu}{\pi\left(4\alpha - 3\right)} &\mbox{if } \alpha \ge \frac{3}{4},
	\end{cases}
\]
with $C_\alpha$ as defined in \eqref{eq:def_C_alpha}.
\end{theorem}

@@

\begin{corollary}[Asymptotic behavior of local clustering function in $G_{\H,n}(\alpha,\nu)$]
\label{cor:asymptotic_local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $(k_n)_{n \ge 1}$ be any positive sequence such that $k_n \to \infty$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$ and $s_\alpha(k)$ defined as in~\eqref{eq:def_scaling_function}. Then, as $n \to \infty$,
\[
	\frac{c_{\H,n}(k_n)}{s_\alpha(k_n)} \, \stackrel{L^1}{\rightarrow} \, 
	\begin{cases}
				C_\alpha \left(\frac{4\alpha \nu}{\pi\left(2\alpha - 1\right)}\right)^{4\alpha - 2}
				&\mbox{if } \frac{1}{2} < \alpha < \frac{3}{4},\\
				\frac{3\nu}{\pi} &\mbox{if } \alpha = \frac{3}{4},\\
				\frac{8\alpha \nu}{\pi\left(4\alpha - 3\right)} &\mbox{if } \alpha \ge \frac{3}{4},
		\end{cases}
\]
where $\stackrel{L^1}{\to}$ denotes converges in expectation.
\end{corollary}

@@

\paragraph{Local clustering function for growing versus fixed degrees}

>>>>

%\MS{Here, I would just say that $B(a,b)$ denotes the beta function $B^-(x,a,b)$ the lower incomplete beta function etc. and put/refer to the integral representations in the appendix, as already done for the Meijer-G function.}
%\subsubsection{Special functions}
We let $B(a,b)$ denote the beta-function and $B^-(x,a,b)$ the lower incomplete beta-function. We also write $\Gamma(z)$ for the Gamma function, $\Gamma^+(q,z)$ for the incomplete Gamma function and define $\Gamma^\ast(q,z) := \Gamma^+(1 + q, z) + \Gamma^+(q, z)$. Finally, we write $U(a,b,z)$ for the hypergeometric U-function (also called Tricomi's confluent hypergeometric function), which for $a,b,z\in \mathbb{C}$, $b \not \in \mathbb{Z}_{\leq 0}$, $\mathrm{Re}(a), \mathrm{Re}(z) >0$ has the integral representation 
\[
	U(a,b,z) = \frac{1}{\Gamma(a)} \int_0^\infty e^{-zt} t^{a-1} (1+t)^{b-a-1} dt,
\] 
see~\cite[p.255 Equation (2)]{erdelyi1953higher}, and let $\MeijerGnew{m}{\ell}{p}{q}{{\bf a}}{{\bf b}}{z}$ denote Meijer's G-Function~\cite{meijer1946gfunction}, see Appendix~\ref{sec:Meijer_G_functions} for more details.

@@

Our first result shows that the local clustering coefficient of the hyperbolic random graph converges in expectation to a constant $c_\infty$, which we can explicitly write down in terms of the special functions defined above.

\begin{theorem}[Limit for local clustering coefficient in $G_{\H,n}(\alpha,\nu)$] \label{thm:clustering_coefficient_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, and $\xi_{\alpha,\nu} = \frac{4\alpha \nu}{\pi(2\alpha -1)}$. Then,
\[
	\lim_{n \to \infty} \Exp{\left|c_{\H,n} - c_\infty\right|} = 0,
\]
where $c_\infty$ is defined for $\alpha \ne 1$ as
\begin{align*}
	c_\infty 
	&=\frac{2 + 4 \alpha + 13 \alpha^2 - 34 \alpha^3 - 12\alpha^4 + 24 \alpha^5}
		{16(\alpha-1)^2 \alpha (\alpha+1) (2\alpha+1)} 
		+  \frac{2^{-1 - 4 \alpha}}{(\alpha - 1)^2} \\
	&\hspace{10pt}+ \frac{(\alpha - 1/2) (B(2 \alpha, 2 \alpha + 1) + B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha))}
		{2 (\alpha - 1) (3 \alpha - 1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{2\alpha} \Gamma^\ast( - 2 \alpha, \xi_{\alpha,\nu})}{4(\alpha-1)}
		+ \frac{\xi_{\alpha,\nu}^{2\alpha + 2}\alpha (\alpha - 1/2)^2 \Gamma^\ast(- 2 \alpha - 2, \xi_{\alpha,\nu})}
		{2(\alpha-1)^2} \\
	&\hspace{10pt}- \frac{\xi_{\alpha,\nu}^{2\alpha + 1}\alpha (2\alpha - 1) \Gamma^\ast( - 2 \alpha - 1,\xi_{\alpha,\nu})}{(\alpha-1)}
		- \frac{\xi_{\alpha,\nu}^{6\alpha-2}2^{-4\alpha}(3\alpha - 1)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)^2}\\
	&\hspace{10pt}-\frac{\xi_{\alpha,\nu}^{6\alpha - 2}(\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha)\Gamma^\ast( - 6 \alpha + 2, \xi_{\alpha,\nu})}{(\alpha-1)} \\
	&\hspace{10pt}- \frac{e^{-\xi_{\alpha,\nu}} \Gamma(2\alpha+1) 
		\left(U(2\alpha+1,1-2\alpha,\xi_{\alpha,\nu}) + U(2\alpha+1,2-2\alpha,\xi_{\alpha,\nu})\right)}{4(\alpha-1)} \\
	&\hspace{10pt}+ \frac{\xi_{\alpha,\nu}^{6\alpha - 2} \Gamma(2\alpha+1)\left( 	
		\MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+2,0}{\xi_{\alpha,\nu}}
		+ \MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+3,0}{\xi_{\alpha,\nu}}\right)}{4(\alpha-1)},
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty &= \frac{575 - 12 \pi^2}{576} + \frac{\eta^4(7 + \pi^2)\Gamma^\ast(-4, \eta)}{4}\\
	&\hspace{10pt}- \frac{1}{2} \int_0^1 (1 - 4z + 3z^3)\log(1-z)(z + \eta)e^{-\eta/z} \dd z\\
	&\hspace{10pt}- \int_0^1 \Li_2(z)(z^3 + \eta z^2) e^{-\eta/z} \dd z,		
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function\footnote{Note that the integrals in the expression for $c_\infty$ for $\alpha = 1$ exists: for the first one note that $1-4z+3z^2=(1-z)(1-3z)$, so the integrand can be bounded by $C(1-z)\log(1-z)$ on $[0,1)$ for some constant $C$, which can be continued continuously to the compact interval $[0,1]$ by noting that the limit for $z \rightarrow 1$ is zero, so the integrand is bounded on a bounded domain and hence, this integral is finite; for the second integral note that $\Li_2(z)$ is bounded by $\Li_2(1)$ on $[0,1]$, which is a series with well-known finite limit, so again the integrand is bounded on a bounded domain and hence the second integral is also finite.}.
\end{theorem}

@@

\begin{theorem}[Limit for local clustering function in $G_{\H,n}(\alpha, \nu)$]
\label{thm:local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\xi = \frac{4\alpha\nu}{\pi(2\alpha-1)}$ and $(k_n)_{n \ge 1}$ be any positive sequence, possibly constant, such that $k_n \ge 3$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, 
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{c_\infty(k_n)} - 1\right|} = 0,
\]
where the function $c_\infty(k)$ is defined for $\alpha \ne 1$ as
\begin{align*}
c_\infty(k)  =&\frac{1}{8\alpha (\alpha-1)\Gamma^+(k-2\alpha,\xi)} \left( -\Gamma^+(k - 2 \alpha, \xi) - 2\frac{\alpha (\alpha - 1/2)^2 \xi^{2} \Gamma^+(k - 2 \alpha - 2, \xi)}{(\alpha - 1)} \right. \\ 
&\left.+ 8 \alpha (\alpha - 1/2) \xi \Gamma^+(k - 2 \alpha - 1,\xi) \right.\\ 
&\left.+ 4\xi^{4\alpha - 2} \Gamma^+(k - 6 \alpha + 2, 
      \xi) \left( \frac{2^{ - 4\alpha}(3 \alpha - 1)}{(\alpha - 1)} + (\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha) \right)  \right.\\ 
&\left.+ \xi^{k-2\alpha} \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi) \right. \\ 
&\left.- \xi^{4\alpha-2} \Gamma(2\alpha+1)\MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}  \right)
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty(k) &= \frac{9 \eta^3}{2 k!} 	
		\Gamma^+(k-3,\eta)-\frac{\xi_{\alpha,\nu}^4}{k!}\frac{7+\pi^2}{4}\Gamma^+(k-4,\eta)\\
	&\hspace{10pt}+ \frac{\eta^k}{2k!}\int_0^1 (1-4z+3z^2)\ln(1-z)z^{1-k}e^{-\eta/z}\dd z\\ 
	&\hspace{10pt}+ \frac{\eta^k}{k!}\int_0^1 z^{3-k} \Li_2(z) e^{-\eta/z} \dd z,
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function.
\end{theorem}

@@

\begin{theorem}[Asymptotic behavior of local clustering function limit]\label{thm:asymptotics_average_clustering_P}
Let $\alpha > \frac{1}{2}$, $\nu > 0$. Then,
\[
	\lim_{k \to \infty} \frac{c_\infty(k)}{s_\alpha(k)} 
	= \begin{cases}
			C_\alpha \left(\frac{4\alpha \nu}{\pi\left(2\alpha - 1\right)}\right)^{4\alpha - 2}
			&\mbox{if } \frac{1}{2} < \alpha < \frac{3}{4},\\
			\frac{3 \nu}{\pi} &\mbox{if } \alpha = \frac{3}{4},\\
			\frac{8\alpha \nu}{\pi\left(4\alpha - 3\right)} &\mbox{if } \alpha > \frac{3}{4},
	\end{cases}
\]
with $C_\alpha$ as defined in \eqref{eq:def_C_alpha}.
\end{theorem}

@@

\begin{corollary}[Asymptotic behavior of local clustering function in $G_{\H,n}(\alpha,\nu)$]
\label{cor:asymptotic_local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $(k_n)_{n \ge 1}$ be any positive sequence such that $k_n \to \infty$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$ and $s_\alpha(k)$ defined as in~\eqref{eq:def_scaling_function}. Then, as $n \to \infty$,
\[
	\frac{c_{\H,n}(k_n)}{s_\alpha(k_n)} \, \stackrel{L^1}{\rightarrow} \, 
	\begin{cases}
				C_\alpha \left(\frac{4\alpha \nu}{\pi\left(2\alpha - 1\right)}\right)^{4\alpha - 2}
				&\mbox{if } \frac{1}{2} < \alpha < \frac{3}{4},\\
				\frac{3\nu}{\pi} &\mbox{if } \alpha = \frac{3}{4},\\
				\frac{8\alpha \nu}{\pi\left(4\alpha - 3\right)} &\mbox{if } \alpha > \frac{3}{4},
		\end{cases}
\]
where $\stackrel{L^1}{\to}$ denotes convergence in expectation.
\end{corollary}
\MS{Are you sure all constants are correct? I thought for $\alpha = 3/4$ it might be $6\nu/\pi$. But if you are sure, you can remove this comment of course.}

@@

\paragraph{Local clustering function for growing versus fixed degrees}
\MS{I'm currently not sure what we should comment here because now Thm 1.2 does cover both cases and the asymptotics in 1.4 appear as a very natural corollary.

However, if we can, then it would be interesting to remark that there is this phase transition in $\alpha$ which appears when the power-law exponent is $2.5=\frac{5}{2}$ and it would be cool if we could give a brief intuitive reason for that or mention that this phase transition point has also been observed in other contexts/models etc.}
